---
title: "Assignment 1 - Language Development in ASD - part 3"
author: "Riccardo Fusaroli"
date: "August 10, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Welcome to the third exciting part of the Language Development in ASD exercise

In this exercise we will delve more in depth with different practices of model comparison and model selection, by first evaluating your models from last time, then learning how to cross-validate models and finally how to systematically compare models.

N.B. There are several datasets for this exercise, so pay attention to which one you are using!

1. The (training) dataset from last time (the awesome one you produced :-) ).
2. The (test) datasets on which you can test the models from last time:
* Demographic and clinical data: https://www.dropbox.com/s/ra99bdvm6fzay3g/demo_test.csv?dl=0
* Utterance Length data: https://www.dropbox.com/s/uxtqqzl18nwxowq/LU_test.csv?dl=0
* Word data: https://www.dropbox.com/s/1ces4hv8kh0stov/token_test.csv?dl=0

### Exercise 1) Testing model performance

How did your models from last time perform? In this exercise you have to compare the results on the training data () and on the test data. Report both of them. Compare them. Discuss why they are different.

- recreate the models you chose last time (just write the code again and apply it to Assignment2TrainData1.csv)

```{r}
#Loading libraries
library(pacman)
p_load(modelr,lmerTest,ggplot2,dplyr,tidyverse,MuMIn,stringr,plyr,caret,Metrics,groupdata2,pastecs,DrugClust)

#setting working directory
setwd("C:/Users/Ejer/Desktop/3. semester/Experimental Methods 3/Assignment1")

#Loading training data
training_data=read.csv("Assignment_1_data.csv",sep=",")
training_data=na.omit(training_data)

#Model from Assignment 1, part 2 - Our awesome model
MLUmodel=lmerTest::lmer(CHI_MLU~1+Visit*Diagnosis+verbalIQ+(1+Visit|ID),training_data,REML=F)
``` 



- calculate performance of the model on the training data: root mean square error is a good measure. (Tip: google the functions rmse() and predict() )

```{r}
#Model performance on training data
modelr::rmse(MLUmodel,training_data)
```

- create the test dataset (apply the code from assignment 1 part 1 to clean up the 3 test datasets)
```{r}
#Loading test data sets
demotest=read.csv("demo_test.csv")
LUtest=read.csv("LU_test.csv")
tokentest=read.csv("token_test.csv")

#renaming variables
demotest=plyr::rename(demotest, c("Child.ID"="ID"))
LUtest=plyr::rename(LUtest,c("SUBJ"="ID","VISIT"="Visit"))
tokentest=plyr::rename(tokentest,c("SUBJ"="ID","VISIT"="Visit"))

#Extracting "phrases" in Visit - leaving only digits
LUtest$Visit=str_extract(LUtest$Visit,"\\d")
tokentest$Visit=str_extract(tokentest$Visit, "\\d")

#Removing "." in the ID coloumn
LUtest$ID=gsub("\\.","",LUtest$ID)
tokentest$ID=gsub("\\.","",tokentest$ID)
demotest$ID=gsub("\\.","",demotest$ID)

#renaming variables name in demotest set
demotest=plyr::rename(demotest,c("MullenRaw"="nonVerbalIQ", "ExpressiveLangRaw"="verbalIQ"))

#Merging data
NewData=merge(demotest,LUtest)
MergedData=merge(tokentest,NewData)

#Selecting needed variables
RealData=dplyr::select(MergedData, ID, Visit, Ethnicity, Diagnosis, Gender, Age, ADOS,  nonVerbalIQ, verbalIQ, MOT_MLU, MOT_LUstd, CHI_MLU, CHI_LUstd, types_MOT, types_CHI, tokens_MOT, tokens_CHI)

df_temp = subset(RealData, Visit == 1) #choosing only the first visit
df_temp = dplyr::select(df_temp, ID, ADOS, nonVerbalIQ, verbalIQ) #choose the relevant info
RealData = subset(RealData, select=-c(ADOS,nonVerbalIQ, verbalIQ))#drop the column you want to refresh

#Merge into the newest and coolest dataframe!
test_data = merge(df_temp, RealData)

#turning ID into numbers:
test_data$ID=as.numeric(as.factor(test_data$ID))

#turning Visit into numeric:
test_data$Visit=as.numeric(as.factor(test_data$Visit))

#renaming values in Gender and Diagnosis
test_data$Gender=as.factor(test_data$Gender)
test_data$Gender=plyr::revalue(test_data$Gender,c("1"="F", "2"="M"))
test_data$Diagnosis=plyr::revalue(test_data$Diagnosis, c("A"="ASD","B"="TD"))

```

- test the performance of the models on the test data (Tips: time to reuse "predict()")

```{r}
#Model performance on test data - how good is our model to predict new data?
modelr::rmse(MLUmodel,test_data)

```

- optional: predictions are never certain, can you identify the uncertainty of the predictions? (e.g. google predictinterval())


### Exercise 2) Model Selection via Cross-validation (N.B: ChildMLU!)

One way to reduce bad surprises when testing a model on new data is to train the model via cross-validation. 

In this exercise you have to use cross-validation to calculate the predictive error of your models and use this predictive error to select the best possible model.

- Create the basic model of ChildMLU as a function of Time and Diagnosis (don't forget the random effects!).

```{r}
#Creating basic model
model1=lmerTest::lmer(CHI_MLU~1+Visit+Diagnosis+(1+Visit|ID),training_data,REML=F)
```
- Make a cross-validated version of the model. (Tips: google the function "createFolds";  loop through each fold, train a model on the other folds and test it on the fold)
- Report the results and comment on them.
```{r}

library(groupdata2)

numfolds = 5
Original_train = training_data
training_data = fold(training_data, numfolds ,id_col="ID", cat_col=c("Gender","Diagnosis"))

perf_test_list=c()
perf_train_list=c()

for (fold in seq(numfolds)){
  temp_train = subset(training_data, .folds !=fold)
  temp_test = subset(training_data, .folds == fold)
  temp_model=lmerTest::lmer(CHI_MLU~1+Visit+Diagnosis+(1+Visit|ID),temp_train,REML=F) #training the simple model 
  
  pred_test=predict(temp_model,temp_test, allow.new.levels=T) #getting performance for test data
  perf_test=Metrics::rmse(temp_test$CHI_MLU, pred_test) 
  perf_test_list=c(perf_test_list,perf_test)
  pred_train=predict(temp_model,temp_train)#get performance for training data
  perf_train=Metrics::rmse(temp_train$CHI_MLU,pred_train)
  perf_train_list=c(perf_train_list,perf_train)

}

mean(perf_test_list)
mean(perf_train_list)

std.error(perf_test_list)
std.error(perf_train_list)

test_data$ID=test_data$ID+100



pred_train = predict(model1,training_data,allow.new.levels=T)
train_rmse = Metrics::rmse(training_data$CHI_MLU,pred_train)


```
- Now try to find the best possible predictive model of ChildMLU, that is, the one that produces the best cross-validated results.

```{r}
library(groupdata2)
library(plotrix)

numfolds = 5

perf_test_list=c()
perf_train_list=c()

#Making some random modelsstrings, which we can test against our cool model from Part 2 (model2) - we want to find the best" predivtive model of ChildMLU 
model2="CHI_MLU~1+Visit*Diagnosis+verbalIQ+(1+Visit|ID)"
model3="CHI_MLU~1+Visit+Diagnosis+MOT_MLU+tokens_MOT+types_MOT+(1+Visit|ID)"
model4="CHI_MLU~1+Visit+Diagnosis+tokens_CHI*types_CHI+(1+Visit|ID)"
model5="CHI_MLU~1+Visit+Diagnosis+nonVerbalIQ+(1+Visit|ID)"
models=c(model2,model3,model4,model5)


Cross_val = function(numfolds, model_string, data){
  training_data = fold(data, numfolds ,id_col="ID", cat_col=c("Gender","Diagnosis"))
  
  perf_test_list=c()
  perf_train_list=c()
  
  for (fold in seq(numfolds)){
    temp_train = subset(training_data, .folds !=fold)
    temp_test = subset(training_data, .folds == fold)
    temp_model=lmerTest::lmer(model_string,temp_train,REML=F) #training the simple model 
    
    pred_test=predict(temp_model,temp_test, allow.new.levels=T) #getting performance for test data
    perf_test=Metrics::rmse(temp_test$CHI_MLU, pred_test) 
    perf_test_list=c(perf_test_list,perf_test)
    pred_train=predict(temp_model,temp_train)#get performance for training data
    perf_train=Metrics::rmse(temp_train$CHI_MLU,pred_train)
    perf_train_list=c(perf_train_list,perf_train)
  
  perf_test_list
  perf_train_list
  }
  
return(data.frame(perf_test = perf_test_list, perf_train = perf_train_list, mean_perf_test = mean(perf_test_list), std.error_test = std.error(perf_test_list), mean_perf_train = mean(perf_train_list), std.error_train = std.error(perf_train_list), model = model_string))
}

for (modelString in models){ 
  if (modelString == models[1]){
  result_df = Cross_val(4, modelString, Original_train)
  result_df
  }
  else{
  result_df = rbind(result_df, Cross_val(4, modelString, Original_train))
  }
}
result_df$modelnr = sort(rep(seq(4), 4))


test_data$ID=test_data$ID+100

#pred_test=predict(model1,test_data, allow.new.levels=T)
#Metrics::rmse(test_data$CHI_MLU, pred_test)
```

- Bonus Question 1: How would you go comparing the performance of the basic model and the cross-validated model on the testing set?
- Bonus Question 2: What is the effect of changing the number of folds? Can you plot RMSE as a function of number of folds?
- Bonus Question 3: compare the cross-validated predictive error against the actual predictive error on the test data

### Exercise 3) Assessing the single child

Let's get to business. This new kiddo - Bernie - has entered your clinic. This child has to be assessed according to his group's average and his expected development.

Bernie is one of the six kids in the test dataset, so make sure to extract that child alone for the following analysis.

You want to evaluate:

- how does the child fare in ChildMLU compared to the average TD child at each visit? Define the distance in terms of absolute difference between this Child and the average TD.
(Tip: recreate the equation of the model: Y=Intercept+BetaX1+BetaX2, etc; input the average of the TD group  for each parameter in the model as X1, X2, etc.).

```{r}
#Extracting Bernie's data
bernie_frame=subset(test_data, ID==302)
summary(model1)

#Constructing a dataframe with the predicted value for the avreage TD child - comparing Bernies performance to this value
td_df=data.frame(ID=300,Visit=1:6,Diagnosis="TD")
predict(model1,td_df, allow.new.levels=T)
abs(predict(model1,td_df, allow.new.levels=T)-bernie_frame$CHI_MLU)

```

- how does the child fare compared to the model predictions at Visit 6? Is the child below or above expectations? (tip: use the predict() function on Bernie's data only and compare the prediction with the actual performance of the child)
```{r}
predict(model1,bernie_frame,allow.new.levels=T)
bernie_frame$CHI_MLU-predict(model1,bernie_frame,allow.new.levels=T)
```

[HERE GOES YOUR ANSWER]

```{r}
write.csv(Original_train,"df_train_original.csv")
write.csv(test_data,"df_test_original.csv")
```

### OPTIONAL: Exercise 4) Model Selection via Information Criteria
Another way to reduce the bad surprises when testing a model on new data is to pay close attention to the relative information criteria between the models you are comparing. Let's learn how to do that!

Re-create a selection of possible models explaining ChildMLU (the ones you tested for exercise 2, but now trained on the full dataset and not cross-validated).

Then try to find the best possible predictive model of ChildMLU, that is, the one that produces the lowest information criterion.

- Bonus question for the optional exercise: are information criteria correlated with cross-validated RMSE? That is, if you take AIC for Model 1, Model 2 and Model 3, do they co-vary with their cross-validated RMSE?

### OPTIONAL: Exercise 5): Using Lasso for model selection
Welcome to the last secret exercise. If you have already solved the previous exercises, and still there's not enough for you, you can expand your expertise by learning about penalizations. Check out this tutorial: http://machinelearningmastery.com/penalized-regression-in-r/ and make sure to google what penalization is, with a focus on L1 and L2-norms. Then try them on your data!


