---
title: "Assignment 1 - Language Development in ASD - part 4"
author: "Riccardo Fusaroli"
date: "August 10, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Welcome to the fourth exciting part of the Language Development in ASD exercise

In this exercise we will assess how many participants we would need to adequately replicate our findings (ensuring our sample size is adequate, our alpha at 0.05 and our beta at 0.8).

### Exercise 1

How much power does your study have (if your model estimates are quite right)?
- [GitHub]Load your dataset, fit your favorite model, assess power for your main effects and interactions of interest.

```{r}
#Setting working directory
setwd("C:/Users/Ejer/Desktop/3. semester/Experimental Methods 3/Assignment1")

#Loading libraries
library(pacman)
p_load(modelr,lmerTest,ggplot2,dplyr,tidyverse,MuMIn,stringr,plyr,caret,Metrics,groupdata2,simr,MASS)


#Loading data
train_data=read.csv("df_train_original.csv",sep=",")
train_data$X.1=NULL


test_data2=read.csv("df_test_original.csv",sep=",")
test_data2$Age=as.character(test_data2$Age)
test_data2$Age=as.numeric(gsub(",",".",test_data2$Age))
test_data2$Gender = rep("M",35)
train_data$Gender = mapvalues(train_data$Gender, from = c("M", "F"), to = c("F", "M"))


MLU_df=rbind(train_data,test_data2)

#Fitting our favourite model (CHild Mlu predcited by verbalIQ and an interaction between diagnosis and visit)
MLU_model=lmerTest::lmer(CHI_MLU~1+Visit*Diagnosis+verbalIQ+(1+Visit|ID),MLU_df,REML=F)
#Making a model without interaction to access the power of individual main effects
MLU_model_noIn = lmerTest::lmer(CHI_MLU~1+Visit+Diagnosis+verbalIQ+(1+Visit|ID),MLU_df,REML=F)
summary(MLU_model) #To get beta-estimates
summary(MLU_model_noIn) #To get beta-estimates

#Accessing power of MLU_model main effects usin 100 simulations
power_visit = powerSim(MLU_model_noIn,fixed("Visit"), nsim = 100)
power_visit #Overpowered - we could get away with less participants

power_diagnosis = powerSim(MLU_model_noIn,fixed("Diagnosis"), nsim = 100)
power_diagnosis #Underpowered - we need more particpants

power_verbalIQ = powerSim(MLU_model_noIn,fixed("verbalIQ"), nsim = 100) 
power_verbalIQ

#Accessing power of interaction between Visit and Diagnosis
power_Interaction = powerSim(MLU_model,fixed("Visit:Diagnosis"),nsim=100)
power_Interaction  
 
```

- Report the power analysis and comment on what you can (or cannot) use its estimates for.

### Exercise 2

How would you perform a more conservative power analysis?
- Identify and justify a minimum effect size for each of your relevant effects
- [GitHub] take the model from exercise 1 and replace the effects with the minimum effect size that you'd accept.
- [GitHub] assess the power curve by Child.ID, identifying an ideal number of participants to estimate each effect
- OPTIONAL if your power estimates do not reach an acceptable threshold simulate additional participants and repeat the previous analysis
- Report the power analysis and comment on what you can (or cannot) use its estimates for.

```{r}

#Identify and justify a minimum effect size for relevant effects
summary(MLU_model)
summary(MLU_model_noIn)

#Identify miminum effect size --> use meta-analysis (can be used as an anchor point), pilot studies etc. Look at estimates and standard error. Look at estimate.

#Does dierction of fixed effects matter? It doesn't matter (e.g. you have a minus with DiagnosisTD)

#Identifying and Jusitfying the minimum effect size used! Adding (less conservative) or substracting (more conservative) the SD - to detect a smaller effect size takes more power

#Main effects
fixef(MLU_model_noIn)["Visit"] = 0.234847-0.023241

#Plotting the power for Visit when fixed effect is changed
powerCurve_visit = powerCurve(MLU_model_noIn,fixed("Visit"),along="ID", nsim = 100)
plot(powerCurve_visit)

fixef(MLU_model_noIn)["verbalIQ"] = 0.07541-0.007148
powerCurve_verbalIQ = powerCurve(MLU_model_noIn, fixed("verbalIQ"), along = "ID", nsim = 100)
plot(powerCurve_verbalIQ)

fixef(MLU_model_noIn)["DiagnosisTD"] = 0.170496-0.089315 #Adding the standard error to be less conservative - diagnosis were underpowered; makes no sense to be more conservative
summary(MLU_model_noIn)

#Plotting the power for Diagnosis when fixed effect is changed
powerCurve_Diagnosis = powerCurve(MLU_model_noIn, fixed("Diagnosis"),along = "ID", nsim = 100)
plot(powerCurve_Diagnosis)

summary(MLU_model)

#Interaction efects --> adding and substracting sd
fixef(MLU_model)["Visit:DiagnosisTD"] = 0.248988-0.035399
fixef(MLU_model)["Visit"] = 0.102829-0.025677
fixef(MLU_model)["DiagnosisTD"] = -0.389235+0.120834
fixef(MLU_model)["verbalIQ"] = 0.075723-0.007136


powerCurve_interaction = powerCurve(MLU_model,fixed("Visit:Diagnosis"),along = "ID", nsim = 100)
plot(powerCurve_interaction)

```


### Exercise 3

Assume you have only the resources to collect 30 kids (15 with ASD and 15 TDs). Identify the power for each relevant effect and discuss whether it's worth to run the study and why.

```{r}
#Creating a small dataframe with 15 TD children and 15 ASD children
TD_kids = unique(MLU_df$ID[MLU_df$Diagnosis == "TD"])[1:15]
ASD_kids = unique(MLU_df$ID[MLU_df$Diagnosis == "ASD"])[1:15]
Allthemkids = c(TD_kids,ASD_kids)

small_df = MLU_df[MLU_df$ID %in% Allthemkids,]

#Fitting the our model with the small dataframme
Small_model_noIN = lmerTest::lmer(CHI_MLU~1+Visit+Diagnosis+verbalIQ+(1+Visit|ID),small_df,REML=F)
Small_model = lmerTest::lmer(CHI_MLU~1+Visit*Diagnosis+verbalIQ+(1+Visit|ID),small_df,REML = F)
summary(Small_model_noIN)
summary(Small_model)

#Assessing power for predictors (might not be necessary):
power_visit2 = powerSim(Small_model_noIN,fixed("Visit"), nsim = 100)
power_visit2

power_verbalIQ2 = powerSim(Small_model_noIN,fixed("verbalIQ"), nsim = 100)
power_verbalIQ2

power_diagnosis2 = powerSim(Small_model_noIN,fixed("Diagnosis"), nsim = 100)
power_diagnosis2

power_Inter2 = powerSim(Small_model,fixed("Visit:Diagnosis"), nsim = 100)
power_Inter2

#Producing a powerCurve for each variable to see whether it's worth running the study
powerCurve_visit = powerCurve(Small_model_noIN,fixed("Visit"),along="ID", nsim = 100)
plot(powerCurve_visit)

powerCurve_Diagnosis = powerCurve(Small_model_noIN, fixed("Diagnosis"),along = "ID", nsim = 100)
plot(powerCurve_Diagnosis)

powerCurve_verbalIQ = powerCurve(Small_model_noIN,fixed("verbalIQ"), along ="ID",nsim = 100)
plot(powerCurve_verbalIQ)

powerCurve_interaction = powerCurve(Small_model, fixed("Visit:Diagnosis"),along = "ID", nsim = 100)
plot(powerCurve_interaction)
```



